---
title: "R Notebook"
output:
  html_notebook:
    theme: cosmo
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}

pacman::p_load(plotly, tidyverse, gapminder, here, vembedr)

```

## gapminder example

Let's recreate Hans Rosling's famous visualization. If you have never seen it, check out this long video: <https://www.ted.com/talks/hans_rosling_the_best_stats_you_ve_ever_seen?utm_campaign=tedspread&utm_medium=referral&utm_source=tedcomshare>

However, for the sake of brevity, watch this video now:

```{r}
embed_url("https://www.youtube.com/watch?v=Z8t4k0Q8e8Y")
```

Get the data from `gapminder`

```{r}
data(gapminder, package = "gapminder")
```

Next, let's create a static `ggplot2` visualization. Note that `frame` and `ids` are the aesthetics that we have never used before. This is because they don't belong to `ggplot2`. They come from `plotly`. So, `ggplot2` will simply ignore them.

Usually we map a time variable to `frame` and cross-sectional id variable to `ids`.

```{r}
gg <- ggplot(gapminder, 
             aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) +
  scale_x_log10() +
  labs(x = 'GDP per Capital',
       y = 'Life Expectancy',
       color = 'Continent') +
  scale_color_manual(values = wesanderson::wes_palette("Moonrise3")) +
  theme_minimal() 
```

Now is the time for some `plotly` magic!

```{r fig.width=9}
ggplotly(gg)
```

## Using `plotly`'s own functionality

Above, we used `ggplot2` to build the graphics and then just modified it using `plotly`. However, `plotly` is a powerful library with tons of functionality of its own. Check out this documentation - <https://plotly.com/r/getting-started/>

Let's create a base plot:

```{r out.width='100%'}
base <- gapminder %>%
  plot_ly(x = ~ gdpPercap,
          y = ~ lifeExp,
          size = ~ pop,
          color = ~ continent,
          text = ~ country,
          hoverinfo = "text") %>%
  layout(xaxis = list(type = "log"))
```

Next, modify the base plot and display it:

```{r message=FALSE, warning=FALSE, fig.width=9}
base %>%
  add_markers(frame = ~ year,
              ids = ~ country) %>%
  animation_opts(1000, 
                 easing = "linear", 
                 redraw = FALSE
                 ) %>%
  animation_button( x = 1,
                    xanchor = "right",
                    y = 0,
                    yanchor = "bottom"
                    ) %>%
  animation_slider(currentvalue = list(prefix = "Year: ",
                                       font = list(color = "red")
                                       )
                   )

```

## Cumulative animation

We will now create a progressing time series of Amazon Inc market value. You can call it a worm plot!

Read Amazon's monthly stock returns from the subdirectory "data" in your project folder.

```{r}
amzn <- read_csv(here::here('data', 'amzn_2005_2020.csv')) %>% 
  mutate(mktval = PRC*SHROUT*1000,
         date2 = as.Date(as.character(date), format = '%Y%m%d')) %>% 
  filter(!is.na(mktval)) %>% 
  mutate(id = row_number())

head(amzn)

```

In order to create a worm, `plotly` needs to create multiple data sets capturing the progress. First data set will effectively have only one observation for the first month. The second data set will have 2 observations, and so on. These are all stacked on top of each other to get a large data set.

```{r}
accumulate_by <- function(dat, var) {
  var <- lazyeval::f_eval(var, dat)
  lvls <- plotly:::getLevels(var)
  dats <- lapply(seq_along(lvls), function(x) {
    cbind(dat[var %in% lvls[seq(1, x)], ], frame = lvls[[x]])
  })
  dplyr::bind_rows(dats)
}

amzn2 <- amzn %>% accumulate_by(~id)
head(amzn2)

```

Finally, the worm plot!


```{r fig.width=9}
p <- amzn2 %>%
  plot_ly(
    x = ~ id, 
    y = ~ mktval,
    frame = ~frame, 
    type = 'scatter',
    mode = 'lines'
  ) %>% 
  layout(
    xaxis = list(
      title = "Date",
      zeroline = F
    ),
    yaxis = list(
      title = "Market Value",
      zeroline = F
    )
  ) %>%
  animation_opts(
    frame = 10,
    transition = 0,
    redraw = FALSE
  ) %>%
  animation_slider(
    hide = T
  ) %>%
  animation_button(
    x = 1,
    xanchor = "right",
    y = 0,
    yanchor = "bottom"
  )

p
```

